Grounding the interaction: anchoring situated discourse in everyday human-robot interaction

### 2011-07-17 cover_letter.tex
Commented out the anecdote on light detection error (naming experiment)
This to address reviewer 3 comment and shorter the paper
Added a note in the cover letter regarding the article length
Raquel's comments

\title{Responses to reviewers' comments and list of modifications done\\ on paper SORO226\\
\LARGE{\textbf{Grounding the interaction: anchoring situated discourse in everyday human-robot interaction}}}
\author{S\'everin Lemaignan \and Raquel Ros \and E. Akin Sisbot \and Rachid Alami \and Michael Beetz}

@@

The authors would like to thank the reviewers for their constructive feedback
that helped us to strengthen the paper.

We next answer the reviewers comments and suggestions. Besides, each answer
indicates the changes made in the paper to address the reviewers' concerns. An
important number of language corrections have been made, and the paper has been
proof-read by a native English speaker.

@@

The analysis of dialogue understanding we have followed has been inspired by
Austin's \emph{speech acts}. Thus, we believe it is worthwhile to briefly mention
them in the introduction as a theoretical ground of our work.

However, since we do not directly refer to them in the remainder of the paper,
we refer to Austin's definitions in a footnote.

@@

A reference to the Dora robot has been added in section 2.1. However, at the
time when this work had been submitted for review, neither of the two proposed
papers by the reviewer were available (and in fact, IJCAI is taking place this
week).

@@

In the current approach, the timing aspect is managed during the analysis of
dialogue. When the user makes reference to an object through demonstratives
(e.g. this, that), the DIALOGS module queries the ontology for objects being
pointed at. The assumption for this mechanism to work is that the human is
constantly pointing while the dialog analysis takes place.  Both situation
assessment and dialogue processing chains induce small delays, but these are
roughly the same (below 200ms), and compensate for each other.

@@

\subsection{``p.8, III.B. "creation of imaginary objects...hinted by verbal
assertions..." -- this phenomenon is called presupposition accommodation''}

@@

{\tt GenerateDescription} translates the parsing tree of each nominal group
into statements. A set of rules (like the one for adjectives mentioned 
in footnote, section 5.1) help with this task.

The detailled algorithm is provided in appendix to this letter.  We would be
happy to add it to the paper as well if the editor wants to consider it.

{\tt Ontlogy.Find} and {\tt Ontlogy.CheckEquivalent} are simple, atomic calls to
the ontology server. The first one send a SPARQL query build from the set of 
partial statement build at the {\tt GenerateDescription} step, the second
one check for each candidate pair if {\tt candidate1 owl:sameAs candidate2}.
This check uses the classified knowledge model.

The {\tt Discrimation} algorithm is explained in details in Ros2010. The
reference has been made more explicit.

@@

When sentences like ``the yellow banana is big" are processed, they are
recognized as statements to be added to the ontology because of the state verb ``to be''
({\sc Dialogs} maintains a list of such verbs).

The nominal group is first resolved as usual (``the yellow banana''), and a new statement
is then generated that describe the new feature to be added to the concept.

Please refer to the {\sc GenerateDescription} algorithm in appendix to see how the
new statements are created.

@@

Thanks for pointing out the work that has been done in this research community.
Even if our contribution focuses more on the grounding of concepts from 
multi-modal knowledge source, these contribution are of definitive 
interest for us.

A reference to Zender's work has been added to the \emph{Related work} 
section, along with a reference to Dale and Reiter paper from 1995.

@@

Paragraph added in the section 6 introduction.

@@

The study was conducted with TUM in Munich. This has been corrected in 
the paper as well.

@@

@@

The method called at (\ref{st.history}) consists in looking through sentences 
that have been stored in the conversation history, then extracting their nominal
group in order to retrieve the identifier of the most recently mentioned 
concept that holds the same characteristics as the nominal group that is being 
processed (algorithm~\ref{algo|History}).

\begin{pseudocode}[ruled]{History.GetMatchingGroup}{group}
\label{algo|History}
\PROCEDURE{History.GetMatchingGroup}{group}
\COMMENT{Extract Nominal group from sentences stored in the history}\\
\mathcal{H} \GETS \CALL{History.GetAllNominalGoup}{}\\
\COMMENT{Generate description of the nominal group that is being processed.} \\
\COMMENT{The adjective  "other" is to be removed before calling this routine} \\
	\mathcal{G} \GETS \CALL{GenerateDescription}{group} \\

>>>>

\title{Answers to reviewers' comments and list of modifications done\\ on paper SORO226\\
\LARGE{\textbf{Grounding the interaction: anchoring situated discourse in everyday human-robot interaction}}}
\author{S\'everin Lemaignan \and Raquel Ros \and E. Akin Sisbot \and Rachid Alami \and Michael Beetz}

@@

The authors would like to thank the reviewers for having accepted to review our work and their constructive feedback, which helped us to strengthen the paper.

We next answer the reviewers comments and suggestions. Besides, each answer
indicates the changes made in the paper to address the reviewers' concerns. An
important number of language corrections have been made with the collaboration of a native English speaker.

Please also note that the overall length of the paper has actually 
increased due to the use of the official Springer \LaTeX template.

@@

The analysis of dialogue understanding we have followed has been inspired by
Austin's \emph{speech acts}. Thus, we believe it is worthwhile to briefly mention
them in the introduction as a theoretical ground of our work. However, since we do not directly refer to them in the remainder of the paper,
we refer to Austin's definitions in a footnote.

@@

A reference to the Dora robot has been added in section 2.1. However, at the
time when this work had been submitted for review, neither of the two papers proposed
by the reviewer were available (and in fact, IJCAI is taking place this
week).

@@

In the current approach, the timing aspect is managed during the analysis of
dialogue. When the user makes reference to an object through demonstrative pronouns
(e.g. this, that), the DIALOGS module queries the ontology for objects being
pointed at. The assumption for this mechanism to work is that the human is
constantly pointing while the dialog analysis takes place.  Both situation
assessment and dialogue processing chains induce small delays, but these are
roughly the same (below 200ms), and compensate for each other.

@@

\subsection{``p.8, III.B. creation of imaginary objects...hinted by verbal
assertions... -- this phenomenon is called presupposition accommodation''}

@@

{\tt GenerateDescription} translates the parsing tree of each nominal group
into statements. A set of rules (like the one for adjectives mentioned 
in footnote, section 5.1) help with this task. The detailed algorithm is provided in appendix to this letter.  We would be happy to add it to the paper as well if the editor wants to consider it.

{\tt Ontlogy.Find} and {\tt Ontlogy.CheckEquivalent} are atomic calls to
the ontology server. The first one sends a SPARQL query based on the set of 
partial statements built at the {\tt GenerateDescription} step. The second
one checks if {\tt $\langle$candidate1 owl:sameAs candidate2$\rangle$} for each candidate pair.
This check uses the classified knowledge model.

The {\tt Discrimation} algorithm is explained in detail in Ros2010. The
reference has been made more explicit.

@@

When sentences like ``the yellow banana is big" are processed, they are
recognized as statements to be added to the ontology because of the state verb ``to be''
({\sc Dialogs} maintains a list of such verbs). The nominal group is first resolved as usual (``the yellow banana''), and a new statement is then generated describing the new feature to be added to the concept. Please refer to the {\sc GenerateDescription} algorithm in the appendix to see how the
new statements are created.

@@

Thanks for pointing out the work that has been done in this research community.
Even if our current work focuses more on the grounding of concepts from 
multi-modal knowledge sources, these contributions are of great interest to us for future work.
A reference to Zender's work has been added to the \emph{Related work} 
section, along with a reference to the Dale and Reiter paper from 1995.

@@

A brief description of both platforms has been included in the introduction of Section 5.

@@

The study was conducted with TUM in Munich. This information has been included in the paper.

@@

To avoid increasing the length of the paper while providing more context about this anecdote, we have finally removed it.

@@

The method called at (\ref{st.history}) consists in looking through sentences 
that have been stored in the conversation history and extracting their nominal
groups. From these we retrieve the identifier of the most recently mentioned 
concept that holds the same characteristics as the nominal group that is being 
processed (algorithm~\ref{algo|History}).

\begin{pseudocode}[ruled]{History.GetMatchingGroup}{group}
\label{algo|History}
\PROCEDURE{History.GetMatchingGroup}{group}
\COMMENT{Extract Nominal group from sentences stored in the history}\\
\mathcal{H} \GETS \CALL{History.GetAllNominalGoup}{}\\
\COMMENT{Generate description of the nominal group that is being processed.} \\
\COMMENT{The adjective  ``other" is to be removed before calling this routine} \\
	\mathcal{G} \GETS \CALL{GenerateDescription}{group} \\

### 2011-07-17 paper_knowledge.tex
Raquel's comments

\begin{abstract}
This paper presents how extraction, representation and use of symbolic
knowledge from real-world perception and human-robot verbal and non-verbal
interaction can actually enable a grounded and shared model of the world that
is suitable for later high-level tasks such as dialogue understanding. We show how
the anchoring process itself relies on the situated nature of human-robot
interactions. We present an implementation, including a specialized symbolic
knowledge representation approach based on Description Logics, and case studies
on several robotic platforms that demonstrate these cognitive capabilities.
\end{abstract}

@@

A messy table, covered with cardboard boxes, books, video tapes... Thomas is
moving out and is packing everything with the help of Jido, his robot
(Fig.~\ref{fig|vpt}).

``Jido, give me that'', says Thomas while looking at a box that contains a
video tape. The robot smoothly grasps the tape, and hands it to the him.

@@

Zender et al.~\cite{Zender2009} address the generation of referring expressions
(\emph{GRE},~\cite{Dale1995}) in situated dialogue for topological knowledge.  They consider
both the reference resolution and reference description tasks, and rely on
OWL-DL representation and SPARQL\footnote{{\em SPARQL Protocol and RDF Query
Language}, \url{http://www.w3.org/TR/rdf-sparql-query/}} to extract
\emph{topological contexts} from their knowledge base.

@@

\begin{figure}[!t] 
	\centering
	\includegraphics[width=\columnwidth]{images/reach.jpg} 
	\caption{Different reaching postures for the human: the bottle and the 
	box are reachable, while the teddy bear is not.} 
	\label{fig::reach} 
\end{figure}

@@

To resolve the nominal group \emph{The yellow banana} a set of partial
statements that describe the concept is generated based on the grammatical
parsing of the sentence (algorithm~\ref{algo|Resolution}(\ref{res.desc})). In the example, a banana (\stmt{?obj type Banana})
that is yellow (\stmt{?obj hasColor yellow})\footnote{Predicates like
\concept{hasColor} or \concept{hasSize} that bind \concept{banana\_01} to
adjectives are extracted from a predefined database of $[Predicate \rightarrow
AdjectiveCategory]$, and falls back on the generic \concept{hasFeature}
predicate if the adjective is not known.}.  Based on these partial statements a
query is sent to the ontology server to retrieve possible instances that match
the description (algorithm~\ref{algo|Resolution}(\ref{res.onto})).

@@

This section presents several experiments where the different features of our
approach are represented.

Experiments have been conducted on two different platforms: the
\emph{Rosie} manipulator from the Technical Universtiy of Munich, a dual-arm,
holonomic service robot, running the ROS\footnote{\emph{Robotic Operating
System}, \url{http://www.ros.org}} middleware, and the \emph{Jido} robot~\cite{Alami1998a} from
LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
Genom/pocolibs} stack.

>>>>

\begin{abstract}
This paper presents how extraction, representation and use of symbolic
knowledge from real-world perception and human-robot verbal and non-verbal
interaction can actually enable a grounded and shared model of the world that
is suitable for later high-level tasks such as dialogue understanding. We show how
the anchoring process itself relies on the situated nature of human-robot
interactions. We present an integrated approach, including a specialized symbolic
knowledge representation approach based on Description Logics, and case studies
on several robotic platforms that demonstrate these cognitive capabilities.
\end{abstract}

@@

A messy table, covered with cardboard boxes, books, video tapes... Thomas is
moving out and is packing everything with the help of Jido, his robot
(Fig.~\ref{fig|vpt}). ``Jido, give me that'', says Thomas while looking at a box that contains a
video tape. The robot smoothly grasps the tape, and hands it to the him.

@@

Zender et al.~\cite{Zender2009} address the generation of referring expressions
(\emph{GRE}~\cite{Dale1995}) in situated dialogue for topological knowledge.  They consider
both the reference resolution and reference description tasks, and rely on
OWL-DL representation and SPARQL\footnote{{\em SPARQL Protocol and RDF Query
Language}, \url{http://www.w3.org/TR/rdf-sparql-query/}} to extract
\emph{topological contexts} from their knowledge base.

@@

\begin{figure}[!t] 
	\centering
	\includegraphics[width=0.8\columnwidth]{images/reach.jpg} 
	\caption{Different reaching postures for the human: the bottle and the 
	box are reachable, while the teddy bear is not.} 
	\label{fig::reach} 
\end{figure}

@@

To resolve the nominal group \emph{The yellow banana} a set of partial
statements that describe the concept is generated based on the grammatical
parsing of the sentence (algorithm~\ref{algo|Resolution}(\ref{res.desc})). The parsed tree of each nominal group is translated into statements based on a set of rules.
In the example, a banana (\stmt{?obj type Banana})
that is yellow (\stmt{?obj hasColor yellow})\footnote{Predicates like
\concept{hasColor} or \concept{hasSize} that bind \concept{banana\_01} to
adjectives are extracted from a predefined database of $[Predicate \rightarrow
AdjectiveCategory]$, and falls back on the generic \concept{hasFeature}
predicate if the adjective is not known.}.  Based on these partial statements a
SPARQL query is sent to the ontology server to retrieve possible instances that match
the description (algorithm~\ref{algo|Resolution}(\ref{res.onto})).

@@

This section presents several experiments where the different features of our
approach are represented. The experiments have been conducted on two different platforms: the
\emph{Rosie} manipulator from the Technical Universtiy of Munich, a dual-arm,
holonomic service robot, running the ROS\footnote{\emph{Robotic Operating
System}, \url{http://www.ros.org}} middleware, and the \emph{Jido} robot~\cite{Alami1998a} from LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
Genom/pocolibs} stack. Both platforms share the use of ORO, the ontology server proposed in this work.
